caDDS: Community Assisted Distributed Database for Sequences

### 2020-07-20 Paper2020ACM/MeetingJune25.tex
July 2020 edits

Intro Pg 1
✓ - Remove sentence: [the goal of this intro]
- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
- Intro goal: layoutht background of the work
- Flow: Idea of A lot of data -> Problems -> Proposed Solution
- remove subsections in inntro (do this last)

-Intro Pg 2
✓ - 3,000,000,000 -> 3 billion
✓- Shorten HGP section. Combine HGP + NGS to one paragraph.
✓ - Fix typo: Genomics is is

@@

Overall info guide: 
✓ - tama na yung flow
- lessen info
- compress database to 1 or 2 sentences. Then mention centralized/distributed database.

Into: CAP
- shorten and make it 1 paragraph
- only describe the most important principle and why

Application Architecture: (Move to Theoretical Framework)
- Discuss Client Server in 1 sentence
- Discuss P2P in 1 sentence
- Describe Hybrid 
- First we need to discuss 2 basic architecture: CS, P2P. 
- CS is blank. 
- P2P is blank. 
- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- make the first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Move Dat (move to theoretical framework)
- Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
✓ - FASTA rename to Biological

@@

Objectives:
- specify metrics, activities, na once you do this succeed.
✓- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
✓- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
✓- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
✓ - framework and APIs move to theoretical framework.

@@

Theoretical Framework
- should have a very basic flow
- discusses abstractly anong paglalaruan na ideas
- Discuss Hybrid,CS, P2P in short sentences
- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
- good flow, but too many info, need to shorten it.
- (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

>>>>

✓Intro Pg 1
✓ - Remove sentence: [the goal of this intro]
✓- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
✓- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
✓- Intro goal: layoutht background of the work
✓- Flow: Idea of A lot of data -> Problems -> Proposed Solution
✓- remove subsections in inntro (do this last)

✓Intro Pg 2
✓ - 3,000,000,000 -> 3 billion
✓- Shorten HGP section. Combine HGP + NGS to one paragraph.
✓ - Fix typo: Genomics is is

@@

✓ Overall info guide: 
✓ - tama na yung flow
✓ - lessen info
✓ - compress database to 1 or 2 sentences. Then mention centralized/distributed database.

✓ Into: CAP
✓ - shorten and make it 1 paragraph
✓ - only describe the most important principle and why
✓- CID: Carlll paread through nalang ng nagawa ko

Application Architecture: (Move to Theoretical Framework)
✓ - Discuss Client Server in 1 sentence
✓- Discuss P2P in 1 sentence
✓- Describe Hybrid 
✓- First we need to discuss 2 basic architecture: CS, P2P. 
✓- CS is blank. 
✓- P2P is blank. 
✓- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- make the first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Move Dat (move to theoretical framework)
(Todo monday) - Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
✓ - FASTA rename to Biological

@@

Objectives:
- specify metrics, activities, na once you do this succeed.
✓ - the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
✓ - For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
✓- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
✓ - framework and APIs move to theoretical framework.

@@

Theoretical Framework
✓ - should have a very basic flow
✓ - discusses abstractly anong paglalaruan na ideas
✓ - Discuss Hybrid,CS, P2P in short sentences
✓ - Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
✓ - Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
✓ - good flow, but too many info, need to shorten it.
✓ - (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

### 2020-07-20 Paper2020ACM/Paper2020.tex
July 2020 edits

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Genomic data the past few years have grown exponentially with the technology barely keeping up. The problem faced by genome researchers is the large data set and difficulty transferring the files. Previous distributed databases are either not meant for genome data, or difficult to replicate. This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases. The proposed system has a master node and data nodes. Master node handles administrative work while the Data nodes contain all sequence data. Data nodes should theoretically increase data transfer speeds. While current databases(e.g. NCBI) have more data, more users, and a live platform. The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises. The proposed system is also theoretically faster as long as there are multiple working data nodes. A recommendation is to implement the system and make the code public for the use of other communities.
\end{abstract}

@@

There is a vast amount of information inside the human body. One of these key biological information storage designs are in the genes. These tell each cell how to organize, how to build, maintain the organism\cite{alberts_mole}. These genes inside a single cell (humans) when laid out is 2 meters in length.\cite{ency_sci_tech} And combining all of the cells, the information laid out would reach the moon, and back, 6000 times.

These biological information storage designs or genes are studied in the field of \textit{Genomics}. Genomics studies the whole sets of genes and their interactions \cite[~p.437]{campbell}. The definition by the NIH (National Institute of Health) adds to that definition by including the study of how these genes interact with the environment.\cite{genomics-definition} Genes are large, and the human genome is filled with 3.2 billion pairs of information (or bases) \cite{introgenomics}.

Before these information is analyzed digitally, they must be read into a machine first. This is called \textit{Sequencing}. Here a machine takes fragments of genes and reads them into a digital format. These machines take in a lot of information from different but redundant fragments of genes to get a complete reading. Another set of computational processes are needed to combine these information to form the gene. Back in the 1980s the speed of sequencing was still slow at around 1000 bases a day\cite{campbell}.

In 1993, when sequencing technology was still at its infancy. There was an initiative to complete the entire human genome\cite{introgenomics}. It was a slow process which took 10 years, a lot of collaborations, 3 billion dollars. In 2000 when the human genome project had been ongoing, the speed of sequencing grew to 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}. The technology became faster and more reliable.

@@

The huge data must be stored in a form of digital information storage design as well. These are where the technologies of databases come in. Databases or DBMS (Database Management Systems) manage the digital information, and the access to it. \cite{Silberschatz2010}. The two types of databases are the centralized and the distributed databases\cite{centralizedvsdistributed}. The centralized databases have a single node or a single access point. While distributed databases have multiple nodes, and each node contains a portion or a replica of the data. Distributed databases are more complex but have higher availability, faster response, and are better for collaborative work. For genomic research, when the data storage needs to be increased often, distributed databases are most appropriate.

\subsubsection{CAP theorem for Databases} \label{cap}
There are 3 principles needed when creating a database\cite{Silberschatz2010}. It is given that it is impossible to do all 3 principles at the same time perfectly\cite{gilbert2012perspectives}. The idea is that only two of the three can be focused on at any given time. These principles and limitations need to be taken into consideration when designing the system.

\begin{itemize}
    \item Consistency - Does the data have integrity? To update one part, all other (replicated) parts should be updated quickly as well; Data is accurate and copied faithfully. An example would be that genes for specimen A, when copied to another data node should contain the exact copy of the data, with no changes.
    \item Availability - Is the database usable by clients all of the time? The entire system must be in downtime as infrequently as possible, despite failures.
    \item Partition Tolerance - How resilient is the system when it comes to failures? If one portion breaks, the other portions should still be entirely usable. An example would be a power failure for one system, how would this affect the other systems?
\end{itemize}

Partition tolerance is a de facto requirement of distributed databases and cannot be avoided, so the main trade-off is between availablity and consistency. \cite{Silberschatz2010} Suppose a distributed system prefers consistency over availability. A user updates an SQL row in one partition of the database. Beforehand, this row has already been replicated among many partitions. To ensure that all replicas of the row are consistent, the system must lock this part of the database to prevent other users from reading and editing it, so availability will be negatively affected in the meantime. A distributed system preferring availability, such as a messenger app, will not lock parts of the database in the purpose of being open to serve its users, but cannot guarantee that the data it outputs is always up-to-date. There might be a delay before the data is fully consistent throughout geographic locations.

\subsection{Application Architecture}
Another important consideration is how a system is built. These describe the blueprint and interaction within a database system. These provide insights similar to the questions asked when looking between database types. Database types describe how the data is handled inside the system. But application architecture describes how the users interact with the system. The application architecture is how a networked app is structured across the various end systems (i.e. computers, servers, etc.) \cite{kurose} The end systems are another way to call users or clients. 
The application architecture types fall into either client-server, peer-to-peer, or hybrid architecture.

\subsubsection{Client-Server} \label{client-server}
A client-server architecture relies on a Web server that is always available and has a fixed IP address. When a client (or user) sends a request for data, the server (or DBMS in this example) responds by sending the requested data. The client-server architecture is mainly built to respond to one client at a time. The architecture may handle multiple connections but as these grow bigger, it may not be able to keep up with the increasing number of connections since the server might not be able to handle all the requests.

\subsubsection{P2P} \label{p2p}
A peer-to-peer (P2P) architecture has minimal to no reliance on always-on, dedicated servers. Peers are desktops/laptops (or users) that communicate to each other without passing through a dedicated server (or DBMS). These include internet, BitTorrent, and some other file sharing protocols. An advantage of P2P networks is their scalability, unlike the earlier type. Each user can benefit the whole network by sharing his/her own resources.

However, since P2P doesn't have a central data management system. P2P has its own problems, which are security and incentivizing users to share their resources.

\subsubsection{Hybrid}
A hybrid architecture combines elements of P2P and client-server architectures, attempting to utilize advantages of both.

For a genomic database system, it would be best to utilize Hybrid Application Architecture. Since there needs to be certain features that Client-Server of P2P might not be able to deliver. One would be scalability, another is security and centralized management of the data, which is a feature not provided by P2P.

@@

\subsection{SeqTorr: A distributed scalable database for genomic information}
The focus on SeqTorr was finding a way to scale databases meant specifically for genomic information. There was also a focus on being able to choose information to share and use. The standard genomics workflow consists of uploading and downloading data on an international database like NCBI in Section \ref{NCBI}. Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds, and  institutions can share in the hosting of data.

@@

SeqTorr seems to go in the right direction and focus for genomic databses. But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work difficult.

\subsection{BioTorrents}
BioTorrents focused on P2P and making it easy to share data across biological scientific communities.  Scientific data continues to grow, and so does the demand for easier accessibility. Centralized servers, as stated earlier, using HTTP or FTP cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\cite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later. BitTorrent handles both these problems.

BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol (or P2P). It is much like a public tracker \cite{biotorrents}. The main website hosts .torrent and metadata files. The main website doesn't store data, rather it lets the users keep the data and share it to other users. The main website only facilitates the sharing of these data (via the idea of P2P and torrents).

The main website stores info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets. There is also a torrent file has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously. These torrents act like a directory telling the user who currently has the data for a piece of the file. The system then helps the user download many pieces of the file from many different users.

The client (or user) must have their computer continuously turned on to keep their file available to others. The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users that will keep many torrents available as often as possible.\cite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.

@@

PeerDB is a proposal for a distributed system using P2P. For this to be used in the context of genomic biological work, a main website which facilitates sharing will be need to be put up. PeerDB needs to be installed as well on all the users who will be participating.

@@

\subsection{NCBI} \label{NCBI}
The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}. This is the standard genomic database used by many researches and institutions. We will refer this later on as well as the \textit{base system}. This is the website used to access and store most genomic information\cite{campbell}. It uses a client-server architecture. Users may download data via the graphics interface (website), or may be done via file transfer protocols such as FTP or HTTP \cite{biotorrents}. The latter is preferred since it may be done to automate the download of data.

The SeqTorr authors\cite{seqtorr} also stated that NCBI is the main and current DBMS used to get and share genomic data. But one difficulty using that is updating the current genomic data. There currently is no way to check which data (on your computer) has been updated, which sequences were corrected. And the way the authors solved this was to regularly re-download the entire NCBI database.

@@

\section{Theoretical Framework}

@@

The Dat Protocol is a newly-developed, open source peer-to-peer protocol for sharing data\cite{dat}. With a single command, anyone can start hosting a directory of files for others to download, as long as they have a publicly accessible IP address and port. Other users can download and then host their own copy of the same directory. Downloads are distributed, meaning if a file is shared across multiple active peers, a new user will be downloading that file from those peers, so bandwidth is shared, much like in BitTorrent. A file's data is synced across peers, so any changes made to the original copy of the file will be sent to peers hosting the same file. And like Git, the version history is also stored and made accessible to peers. Unlike Git though, Dat works well for any kind of file, not just source code. Dat also has security built in. Only users with the directory's read key can download it, and only users with the write key can make changes. In short, Dat has "taken the best parts of Git, BitTorrent, and Dropbox". Another important feature is that files and directories are stored as-is, meaning no modifications are done by Dat when storing files.

@@

\subsection{Networks}

An important aspect of the proposed system is the transfer of files to other computers. To achieve that, there are network technologies and systems that need to be understood. Each layer of the network has certain features, securities, and abstraction which contributes to the overall idea of sharing files across networks. Networks Layers are the abstraction of how computers pass data to each other. Understanding this is needed to maximize the speed in which data is transferred between computers. As well as to realize where certain problems or bottlenecks later on. A network, specifically the Internet, is comprised of the following levels of abstraction. \cite{kurose}
\begin{enumerate}
    \item Level 1: Medium used to send signals (physical, electrical signals of 0
and 1), through wire or by air
    \item Level 2: Introduction of packets: WIFI, Ethernet, connecting together
different interfaces. Goal of sending a group of signals.
    \item Level 3: Introduction of networks: IP Address to uniquely identify
entities.
    \item Level 4: Management of error free data transmission: Sending a group
of messages via TCP, UDP, FTP (File Transfer Protocols)
    \item Level 5: Abstraction of Level 4: Function calls programmers will use
to do level 4 processes.
\end{enumerate}

Level 1 through 3 focus on building a system for sending messages from one computer entity to another. Every level talks to another entity on the same level. And a higher level talks to the lower level. This later on will be important to understanding the scope of what the thesis will do. Level 1 and 2 can be vulnerable if anyone can read the data inside the network, partly why HTTPS was developed to encrypt the data.

Level 4 and 5 is what the system will mainly use and operate on. In these systems, there are APIs commonly used to transfer files between computers on networks. These levels, compared to the levels earlier discussed also are more specific to the ideas of file transfer, checking if the files are error free, and more. Making sure files transferred are error free will be needed to do data synchronization later on. Another important network concept is that computers closer on the network (when talking directly) should have a faster speed and lower latency than computers farther apart on a network.

In networks, when there is data being accessed over the network, it is not called C.R.U.D. (the way databases access data), instead it is called POST, GET, PUT, DELETE. POST is analogous to CREATE meaning the user submit data to be stored (in the server). GET is similar to READ but is in the context of reading a file over a network (from a server). PUT is analogous to UPDATE as DELETE remains as DELETE. These are codes used to denote network commands that do the same thing as the CRUD functions discussed earlier. This is important since there is an added layer of difficulty when it comes to transmitting data over a network.

\begin{itemize}
    \item (Move to Theoretical Framework/caDDS) The system is an information system, and all information systems must have the main API’s 
    \begin{itemize}
        \item POST genome data
        \item GET genome data (this will implement a hybrid P2P transfer)
        \begin{itemize}
            \item This will aim to speed up number 3 in the Problems of Classical Databases
        \end{itemize}
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
\end{itemize}

@@

\subsubsection{Overview}
The two parts of the system infrastructure are the master node and data nodes. These are also inspired from SeqTorr.

@@

The proposed method introduces the idea of data nodes and in section \ref{theoreticalspeed} explains how having at least one data node should theoretically rival the speed of the base method. And having multiple data nodes will be faster in theory than the base method. Another advantage of having a distributed database is the community aspect (see Section \ref{community}). The proposed method may be implemented on a local scale, meaning it can have the necessary data for the community. The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}). Another is that the features on the designs of the speed, consistency, and fault tolerance (see Section \ref{cap} ) may be calibrated to the needs of the community.

\section{Conclusion}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Genomics data the past few years have grown exponentially with the technology barely keeping up. The problem faced by genome researchers is the large data set and difficulty transferring the files. Previous distributed databases are either not meant for genome data, or difficult to replicate. This paper lays the groundwork for a distributed database that is designed to easily scale to accommodate bigger storage, and also reduces the data speed bottleneck faced by centralized databases. The proposed system has a master node and data nodes. The master node handles administrative work while the Data nodes contain all sequence data. Data nodes should theoretically increase data transfer speeds. While current databases(e.g. NCBI) have more data, more users, and a live platform. The proposed system is aimed towards a smaller community, with more frequent and localized data, which the data storage can easily be expanded as the need arises. The proposed system is also theoretically faster as long as there are multiple working data nodes. A recommendation is to implement the system and make the code public for the use of other communities.
\end{abstract}

@@

There is a vast amount of information inside the human body. One of these key biological information storage designs are in the genes. These tell each cell how to organize, how to build, and how to maintain the organism\cite{alberts_mole}. The genes inside a single cell (humans) when laid out is 2 meters in length.\cite{ency_sci_tech} And combining all of the cells, the information laid out would reach the moon, and back, 6000 times.

These biological information storage designs or genes are studied in the field of \textit{Genomics}. Genomics studies the whole sets of genes and their interactions \cite[~p.437]{campbell}. The definition by the NIH (National Institute of Health) adds to that definition by including the study of how these genes interact with the environment.\cite{genomics-definition} Genes are large, the human genome specifically is filled with 3.2 billion pairs of information (or bases) \cite{introgenomics}.

Before these information are analyzed digitally, they must be read into a machine first. This is called \textit{Sequencing}. Here a machine takes fragments of genes and reads them into a digital format. These machines take in a lot of information from different but redundant fragments of genes to get a complete reading. Another set of computational processes are needed to combine these information to form the gene. Back in the 1980s the speed of sequencing was still slow and the top laboratory could only process around 1000 bases a day\cite{campbell}.

In 1993, when sequencing technology was still at its infancy. There was an initiative to complete the entire human genome\cite{introgenomics}. It was a slow process which took 10 years, a lot of collaborations, and 3 billion dollars. In 2000, when the human genome project had been ongoing, the speed of sequencing grew to processing 1000 base pairs every second, 24 hours a day, 7 days a week \cite{campbell}. The technology became faster and more reliable.

@@

The huge data must be stored in a form of digital information storage design as well. These are where the technologies of databases come in. Databases or DBMS (Database Management Systems) manage the digital information, and the access to it. \cite{Silberschatz2010}. The two types of databases are the centralized and the distributed databases\cite{centralizedvsdistributed}. The centralized databases have a single node or a single access point. While distributed databases have multiple nodes, and each node contains a portion or a replica of the data. Distributed databases are more complex but have higher availability, faster response, and are better for collaborative work. For genomics research, when the data storage needs to be increased often, distributed databases are most appropriate.

Databases also have three principles \cite{Silberschatz2010}, these 3 principles form the ideal database. But these 3 principles, practically cannot be done all at the same time \cite{Silberschatz2010}. These are the considerations and limitations that need to be addressed. One principle is consistency, data should be accurate and faithfully replicated. Another is availability, is the database able to serve users all of the time? The down times should be as infrequent as possible. The last and which will be the focus is Partition Tolerance. Partition Tolerance is the resiliency of a system. What happens when one portion malfunctions, will the whole system fail or will the rest easily adjust? This is necessary for a distributed database. And since Partition Tolerance \cite{Silberschatz2010} is chosen, then the trade off will be between availability and consistency.

The trade off for this system is between availability and consistency. For a consistency-priority-database, a change in one portion of the database, a create SQL command, will lock the entire database preventing other changes (reducing availability), then it will make sure that change is replicated in all relevant portions of the database, before making the database available again.

For an availability-priority-database, an example would be the messenger Facebook app. Where the focus of the app is speed. Here changes in a database will not lock the database, to make its app available. The downside is that the data is available, but it is not always up to date, and time is needed before the data is fully consistent on all the relevant databases.

@@

\subsection{SeqTorr: A distributed scalable database for genomic information}
The focus on SeqTorr was finding a way to scale databases meant specifically for genomics information. There was also a focus on being able to choose information to share and use. The standard genomics workflow consists of uploading and downloading data on an international database like NCBI in Section \ref{NCBI}. Having a distributed scalable local infrastructure to store genomic data instead of relying on NCBI would be beneficial since: only certain sequences in NCBI are relevant to the institutions (e.g. Asian sequences are needed more than Caucasian ones), researchers within the country can share data with fast speeds,and institutions can share in the hosting of data.

@@

SeqTorr seems to go in the right direction to focus on genomics databases. But currently, SeqTorr is not available, and the code base does not have any documentation which makes future work on this difficult.

\subsection{BioTorrents}
BioTorrents focused on P2P and making it easy to share data across biological scientific communities. Scientific data continues to grow, and so does the demand for easier accessibility. Centralized servers, as stated earlier, which are using HTTP or FTP, cannot keep up with concurrent requests, and peer-to-peer protocols does not scale well for large files.\cite{biotorrents} One example of these centralized servers is NCBI, which will be discussed later. BitTorrent handles both these problems.

BioTorrents is a system for legally sharing scientific data that works on top of the BitTorrent protocol (or P2P). It is much like a public tracker \cite{biotorrents}. The main website hosts torrent and metadata files. The main website doesn't store data, rather it lets the users keep the data and share it to other users. The main website only facilitates the sharing of these data (via the idea of P2P and torrents).

The main website stores info about each dataset includes categories, license, filenames, etc. that helps users in searching for relevant datasets. There is also a torrent file that has a list of trackers, servers that "know" which peers are serving which files, so it can download from those peers simultaneously. These torrents act like a directory telling the user who currently has the data for a piece of the file. The system then helps the user download many pieces of the file from many different users.

The client (or user) must have their computer continuously turned on to keep their file available to others. The main issue with BioTorrents, which is a problem as well in P2P, is that it requires lots of users to keep torrents available as often as possible.\cite{biotorrents} These entail that for the system to work, many people need to use it and all these people need to keep their computers on all the time.

@@

PeerDB is a proposal for a distributed system using P2P. For this to be used in the context of genomics biological work, a main website which facilitates sharing will be need to be put up. PeerDB needs to be installed as well on all the users who will be participating.

@@

\subsection{NCBI} \label{NCBI}
The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}, is the standard genomics database used by many researches and institutions. We will refer a portion of this later on as the \textit{base system}. This is the website used to access and store most genomics information around the world\cite{campbell}. It uses a client-server architecture. Users may download data via the graphics interface (website), or may be done via file transfer protocols such as FTP or HTTP \cite{biotorrents}. The latter is preferred since it may be done to automate the download of data.

The SeqTorr authors\cite{seqtorr} also stated that NCBI is the main and current DBMS used to get and share genomics data. But one difficulty using that is updating the current genomics dataset. There is currently no way to check which data (on your computer) have been updated, which sequences were corrected. And the way the authors solved this was to regularly re-download the entire NCBI database.

@@

\section{Theoretical Framework}
Before the system features are discussed, notes about the design, and how it connects internally is needed to gain a better appreciation of the overall system

\subsection{Application Architecture}
The system's design is guided by the application architecture. Here it is discussed how an app that relies on networks functions and interacts \cite{kurose}. There are three main types: Client-Server, Peer-to-Peer, and Hybrid. Client-Server or CS means there is one web application that serves the data to multiple users (or clients, hence the name). As the number of clients grow bigger, the server takes more stress serving all the data. Peer-to-Peer or P2P is different since it has no singular dedicated server. No singular point of stress, and data is easily distributed. The term client and server is merged into a term "peer" that can both receive and send data. All peers can share data to other peers but the catch is that there is no centralized authority moderating these data, and that an active user base or peer base is needed for this to function.

Hybrid tackles some of the issues of CS and P2P, which is scalability and moderation. It allows easier scaling of database access (akin to adding more servers in P2P), but it also allows to have a form of moderation and security (akin to adding a moderator to CS). This is also more appropriate for a distributed database, and which is why a hybrid architecture is used.

@@

The Dat Protocol is a newly-developed, open source peer-to-peer protocol for sharing data\cite{dat}. With a single command, anyone can start hosting a directory of files for others to download, as long as they have a publicly accessible IP address and port. Other users can download and then host their own copy of the same directory. Downloads are distributed, meaning if a file is shared across multiple active peers, a new user will be downloading that file from those peers, so bandwidth is shared, much like in BitTorrent. A file's data is synced across peers, so any changes made to the original copy of the file will be sent to peers hosting the same file. And like Git, the version history is also stored and made accessible to peers. Unlike Git though, Dat works well for any kind of file, not just source code. Dat also has security built in. Only users with the directory's read key can download it, and only users with the write key can make changes. In short, Dat has "taken the best parts of Git, BitTorrent, and Dropbox"\cite{dat}. Another important feature is that files and directories are stored as-is, meaning no modifications are done by Dat when storing files.

@@

\subsection{Networks}
In each architecture it was discussed that there is a form of file transfer between peers or from servers to clients. To achieve this transfer, years of technology created systems called networks. There are 5 levels in a network \cite{kurose} but only the last 2, level 4 and 5 is needed for the system. The other levels 1-3 create protocols that send messages to specific and unique computers without error and utilizing any erratic medium between them (whether it be undersea cables or air). 
Level 4 assumes sending between computers is possible, but erratic. And these protocols in this level make sure all data is complete, error free before it is acknowledged. An important feature here is the File Transfer Protocol which is used by the NCBI. Here a client needs to set up a connection to the main NCBI server, then request specific files to be transferred to the client(or the user), and of course to move these files without error.

Level 5 abstracts level 4, and making the process of connecting and transferring easier to accomplish. Since Level 5 uses function calls to do the protocols of level 4. A designer of a system will want the ease-of-use of using level 5 functions, but will need knowledge of level 4 in order to better understand bugs and problems later on. Level 5 also allows the user to focus more on doing similar database or data manipulation functions (C.R.U.D.) as POST, GET, PUT, DELETE.

All information systems must have POST, GET, PUT, DELETE or the corollary (C.R.U.D.). Since there are networks in this system, the network version of the data manipulation functions will be used.

\begin{itemize}
    \item The system being proposed is an information system, and this system will have the main API’s
    \begin{itemize}
        \item POST genome data
        \item GET genome data, this system will implement a hybrid P2P transfer
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
\end{itemize}

@@

\subsubsection{Overview}
The two parts of the system infrastructure are the master node and data nodes. This system is also inspired by SeqTorr.

@@

The proposed method introduces the idea of data nodes and in section \ref{theoreticalspeed} explains how having at least one data node should theoretically rival the speed of the base method. And having multiple data nodes will be faster in theory than the base method. Another advantage of having a distributed database is the community aspect (see Section \ref{community}). The proposed method may be implemented on a local scale, meaning it can have the necessary data for the community. The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}). Another is that the features on the designs of the speed, consistency, and fault tolerance  may be calibrated to the needs of the community.

\section{Conclusion}
The research designed a theoretical design for a database for genomics researchers. The database is distributed and uses a hybrid architecture when managing the data. The theoretical properties of the design was studied and compared to a base method (FTP of NCBI), notably the speed and correctness. The proposed system, is better suited for database systems that scale. And is shown to be faster if there are multiple data nodes.
